Microclimate Variable Correlations

### 2016-09-01 microclimate-correlations.Rmd
Revise variables in correlations and add bivariate plots.

microclimate_variables <- oakplots.sub %>% 
      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, contains("hrs"), contains("tmax"), contains("tmin"))
# summary(microclimate_variables)
```

@@

```{r microclimate correlations, echo=F}
pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2005 Sample Year \n n=165 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2006), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2006 Sample Year \n n=167 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2007), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2007 Sample Year \n n=164 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2008), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2008 Sample Year \n n=163 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2009), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2009 Sample Year \n n=162 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2010), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2010 Sample Year \n n=159 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2011), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2011 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2012), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2012 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2014), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmin_rs, avg_tmax_ds, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2014 Sample Year \n n=156 oak plots")
pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)), 
      lower.panel = panel.cor,
      main = "Correlations Between Variables \n all oak plots 2005-2014")

@@

@@

```{r bivariate models using umca plots, echo=F, fig.height=8, message=F}
umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
umca_plots <- left_join(umca_plots, wet_hrs_lag, by = c("plotid","sample_year"))
umca_plots$tot_lfct.log <- log(umca_plots$tot_lfct+1)
umca_plots$avg_lfct.log <- log(umca_plots$avg_lfct+1)

@@

>>>>

microclimate_variables <- oakplots.sub %>%
      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, inf_oak_ct, uninf_oak_ct, contains("hrs"), contains("tmax"), contains("tmin")) %>% 
      mutate(inf_oak.pct = inf_oak_ct/(inf_oak_ct + uninf_oak_ct))
microclimate_variables <- na.omit(microclimate_variables)
# summary(microclimate_variables)
```

@@

```{r microclimate correlations, echo=F, fig.width=12, fig.height=8, eval=T}
pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2005 Sample Year \n n=165 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2006), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2006 Sample Year \n n=167 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2007), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2007 Sample Year \n n=164 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2008), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2008 Sample Year \n n=163 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2009), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2009 Sample Year \n n=162 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2010), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2010 Sample Year \n n=159 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2011), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2011 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2012), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2012 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2014), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2014 Sample Year \n n=156 oak plots")
pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1)), 
      lower.panel = panel.cor,
      main = "Correlations Between Variables \n all oak plots 2005-2014")

# pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), inf_oak_ct, tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw14_wet_tminus1, hrsblw10_rs_tminus1, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet, avg_hrsblw14_wet_tminus1, avg_hrsblw10_rs_tminus1)), 
#       lower.panel = panel.cor,
#       main = "Correlations Between Variables \n all oak plots 2005-2014")

@@

```{r bivariate linear model hrsblw10rs, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs_blw10rs
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs_blw10rs, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Rainy-season Hours <10 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("Oak Plots") +
      theme_bw()
```

```{r bivariate linear model hrsblw14rs, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs_blw14rs
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs_blw14rs, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Rainy-season Hours <14 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("Oak Plots") +
      theme_bw()
```

```{r bivariate linear model hrsblw14_wet, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrsblw14_wet
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrsblw14_wet, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Wet-hours <14 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("Oak Plots") +
      theme_bw()
```

```{r bivariate linear model hrs1020_wet, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs1020_wet
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs1020_wet, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Wet-hours 10-20 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("Oak Plots") +
      theme_bw()
```

@@

```{r bivariate models using umca plots, echo=F, fig.height=8, message=F}
# umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
umca_plots <- left_join(umca_plots, wet_hrs, by = c("plotid","sample_year"))
umca_plots$tot_lfct.log <- log(umca_plots$tot_lfct+1)
umca_plots$avg_lfct.log <- log(umca_plots$avg_lfct+1)

@@

```{r bivariate model umca plots hrsblw10rs, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs_blw10rs
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                         list(a = format(a, digits = 2),
                              b = format(b, digits = 2),
                              r2 = format(r2, digits = 3)))
      c(lab = as.character(as.expression(expr)))
}
# y <- "avg_hrs1422_wet_t_t1.log"
# x <- "avgtmax_wet"
# reg_eqn(df = na.omit(umca_plots), y = y, x = x)

eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)

ggplot(filter(umca_plots, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs_blw10rs, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Rainy-season Hours <10 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("UMCA Plots") +
      theme_bw()
```

```{r bivariate model umca plots hrsblw14rs, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs_blw14rs
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                         list(a = format(a, digits = 2),
                              b = format(b, digits = 2),
                              r2 = format(r2, digits = 3)))
      c(lab = as.character(as.expression(expr)))
}
# y <- "avg_hrs1422_wet_t_t1.log"
# x <- "avgtmax_wet"
# reg_eqn(df = na.omit(umca_plots), y = y, x = x)

eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)

ggplot(filter(umca_plots, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs_blw14rs, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Rainy-season Hours <14 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("UMCA Plots") +
      theme_bw()
```

```{r bivariate model umca plots hrsblw14_wet, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrsblw14_wet
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                         list(a = format(a, digits = 2),
                              b = format(b, digits = 2),
                              r2 = format(r2, digits = 3)))
      c(lab = as.character(as.expression(expr)))
}
# y <- "avg_hrs1422_wet_t_t1.log"
# x <- "avgtmax_wet"
# reg_eqn(df = na.omit(umca_plots), y = y, x = x)

eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)

ggplot(filter(umca_plots, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrsblw14_wet, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Wet-hours <14 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("UMCA Plots") +
      theme_bw()
```

```{r bivariate model umca hrs1020_wet, echo=F, fig.height=8, message=F}
reg_eqn <- function(df,y,x) {
      y = df$tot_lfct.log
      x = df$hrs1020_wet
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                         list(a = format(a, digits = 2),
                              b = format(b, digits = 2),
                              r2 = format(r2, digits = 3)))
      c(lab = as.character(as.expression(expr)))
}
# y <- "avg_hrs1422_wet_t_t1.log"
# x <- "avgtmax_wet"
# reg_eqn(df = na.omit(umca_plots), y = y, x = x)

eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)

ggplot(filter(umca_plots, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(x = hrs1020_wet, y = tot_lfct.log)) +
      geom_point() +
      geom_smooth(method = "lm", size = 0.7) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
                hjust = 0, vjust = 1) +
      facet_grid(sample_year ~ .) +
      xlab("Wet-hours 10-20 C") +
      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
      ggtitle("UMCA Plots") +
      theme_bw()
```
